NURM: a tool to locate numerical differences in pipelines

### 2019-10-03 paper.tex
Updates

Since all the result files are captured in Condition 1, the tool 
is ready to label the piepeline processes based on the results in Condition 2.
However, another challenge is to label undetermined processes that read files with 
differences and write files with differences too. In this
situation, it is not possible to determine from the pipeline results 
whether the process created differences or 
was transparent.
To address this issue, we set hooks in the pipeline 
in Condition 2. Hooks are set by replacing the pipeline 
processes with a custom script. This custom script labels each process 
as \emph{transparent} or \emph{create differences} immediately after running 
the original executable. 
The replacement is done through the PATH variable, as before. Moreover, 
this custom script copies the results obtained from the same process in 
Condition 1 to the pipeline output in Condition 2 if 
it is invoked with the arguments of a process that created differences 
(Figure~\ref{fig:overview-labelling} - (3)).

Differences are identified by comparing checksum of the output files for 
the same process in both conditions (Figure~\ref{fig:overview-labelling} - (2)).
In addition, the distance functions that are specific to file types can be added 
to our tool to ensure that differences belong to the file data part.
We developed an incremental approach that consists of the following steps:

@@

\subsection{Processing of data}

@@

In our experiments, we identified two processes that create differences due to 
within-OS differences.  
Registration process with \emph{mri\_segreg} and process \emph{mris\_curvature} that 
computes mean and Gaussian curvatures of a cortical surface using 
the random variable in FreeSurfer.
To avoid such pseudo-randomness, we 
initialized random number generators using a seed state in 
the command line in which pipeline script calls the process.
In this way, NURM-tool identifies the processes in the pipelines 
that create between-OS differences.

@@

Table~\ref{table:data-clusters} shows the PreFreeSurfer result of subject clustering. 
Among our dataset, we identified four types of subjects with 
the different topology of \emph{process trees}. 
Differences correspond to the data processing that leads to producing a disparate output file(s).
For example, subject type 3 missed output file \emph{T1w2\_gdc.nii.gz} after processing 
in comparison with subject type 1.
Besides, we verified that 
the \emph{process trees} were identical for all subjects of the same type in 
different versions of CentOS.

\begin{table}
\centering
\begin{threeparttable}
\caption{Clustering of 20 input subjects.}

\begin{tabular}{@{}llllll@{}}
\toprule
Type   & Number of Subjects  & Missing File(s)                                    \\ \midrule
type1  & 9                   & None                                               \\
type2  & 7                   & [``T1w/T1w2\_gdc.nii.gz",                          \\
       &                     &  ``T2w/T2w2\_gdc.nii.gz"]                          \\
type3  & 2                   & [``T1w/T1w2\_gdc.nii.gz"]                          \\
type4  & 2                   & [``T2w/T2w2\_gdc.nii.gz"]                          \\ \bottomrule
\end{tabular}
\begin{tablenotes}
     \small
     \item *Type refers to the topology type of the \emph{process tree} of subjects.
\end{tablenotes}
\end{threeparttable}
\label{table:data-clusters}
\end{table}

The frequency of the processes that create differences in 
the pipeline among 20 subjects is determined in Figure~\ref{fig:pfs_freq}. 
\emph{FLIRT} is the most frequent process that produces differences.
Furthermore, the percentage of processes that introduce differences 
along with the 
sub-pipelines are depicted in this figure 
including linear registration with “\emph{FLIRT}” (in 
ACPC-Alignment, Brain Extraction, Distortion Correction, 
Atlas Registration), non-linear registration with “\emph{FNIRT}” (in 
Brain Extraction and Altas Registration), image warping with 
“\emph{new\_invwarp}” (in Brain Extraction and Atlas Registration). 
Besides, differences were observed in image mean and standard-deviation 
computations in image interpolation with “\emph{fslmaths}” (in Anatomical Average). 
Results show that \emph{fslmaths} and \emph{FLIRT} 
are the processes that at least have one subprocess to produce differences across all the subjects.

>>>>

Since all the result files are captured in Condition 1, the tool 
is ready to label the piepeline processes based on the results in Condition 2.
However, another challenge is to label undetermined processes that read files with 
differences and write files with differences too. In this
situation, it is not possible to determine from the pipeline results 
whether the process created differences or 
was transparent.
To address this issue, we set hooks in the pipeline 
in Condition 2. Hooks are set by replacing the pipeline 
processes with a custom script. This custom script labels each process 
as \emph{transparent} or \emph{create differences} immediately after running 
the original executable. 
The replacement is done through the PATH variable, as before. Moreover, 
this custom script copies the results obtained from the same process in 
Condition 1 to the pipeline output in Condition 2 if 
it is invoked with the arguments of a process that created differences 
(Figure~\ref{fig:overview-labelling} - (3)). 
Differences are identified by comparing checksum of the output files for 
the same process in both conditions (Figure~\ref{fig:overview-labelling} - (2)).
In addition, the distance functions that are specific to file types can be added 
to our tool to ensure that differences belong to the file data part.
We developed an incremental approach that consists of the following steps:

@@

\subsection{Data processing}

@@

We didn't observe any within-OS difference in PreFreeSurfer, however, we
 identified two FreeSurfer processes leading to within-OS differences due
 to the use of pseudo-random numbers: image registration with
 \emph{mri\_segreg}, and cortical surface curvature
 estimations with \emph{mris\_curvature}. To address this issue, we fixed
 the seed used by the pseudo-random number generator, through an option of
 the FreeSurfer tools. In this way, our analysis can focus on between-OS
 differences.

@@

We identified four types of subjects with different PreFreeSurfer process
trees (Table~\ref{table:data-clusters}). Differences between subject types
are coming from different  numbers of T1 and T2 images in the raw data:
subjects of type 1 have 2 T1 images and 2 T2 images, while subjects of type
4 only have 1 T1 image and 1 T2 image. \tristan{check that} This leads to
different sets of output files. We verified that the process trees were
identical for all subjects of the same type in different versions of
CentOS.

\begin{table}
\label{table:data-clusters}
\centering
\begin{threeparttable}
\caption{Clustering of 20 input subjects.}

\begin{tabular}{@{}llllll@{}}
\toprule
Type   & Number of Subjects  & Missing File(s)                                    \\ \midrule
type1  & 9                   & None                                               \\
type2  & 7                   & [``T1w/T1w2\_gdc.nii.gz",                          \\
       &                     &  ``T2w/T2w2\_gdc.nii.gz"]                          \\
type3  & 2                   & [``T1w/T1w2\_gdc.nii.gz"]                          \\
type4  & 2                   & [``T2w/T2w2\_gdc.nii.gz"]                          \\ \bottomrule
\end{tabular}
\begin{tablenotes}
     \small
     \item *Type refers to the topology type of the \emph{process tree} of subjects.
\end{tablenotes}
\end{threeparttable}
\end{table}

The frequency of the processes that create differences in 
the pipeline is shown in Figure~\ref{fig:pfs_freq}. 
\emph{FLIRT} is the most frequent process that produces differences.
Furthermore, the percentage of processes that introduce differences 
along with the 
sub-pipelines are depicted in this figure 
including linear registration with “\emph{FLIRT}” (in 
ACPC-Alignment, Brain Extraction, Distortion Correction, 
Atlas Registration), non-linear registration with “\emph{FNIRT}” (in 
Brain Extraction and Altas Registration), image warping with 
“\emph{new\_invwarp}” (in Brain Extraction and Atlas Registration). 
Besides, differences were observed in image mean and standard-deviation 
computations in image interpolation with “\emph{fslmaths}” (in Anatomical Average). 
Results show that \emph{fslmaths} and \emph{FLIRT} 
are the processes that at least have one subprocess to produce differences across all the subjects.
